Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem

### 2015-12-07 acta_manuscript.tex
finished updating the text. need to generate new pretty figures

%---------------------------- EXPLICIT SUMMARY OF CONTRIBUTIONS--------------------------%
In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
Previous results in the design of optimal transfers have relied on suboptimal direct optimization methods and conventional integration techniques.
This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
The use of a geometric integrators ensures numerical stability for long-duration orbit transfers.
A numerical example is presented demonstrating a transfer trajectory from \( L_1 \) to the vicinity of the Moon and compared to the control-free design methodology.
\begin{itemize}
	\item generalize invariant manifold transfers with continuous control input
	\item Low thrust propulsion rather than impulsive manuevers
	\item Combination of reachable set and \Poincare section
	\item Geometric integrators for accurate numerical simulation
\end{itemize}

@@

% motivation - may need to compute several reachability sets

% discuss difficulties in applying direct optimization or the invariant manifold method

% describe initial and target states for this example
This example demonstrates the feasibility of iteratively computing the reachable set. 
In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
   \caption{Transfer target}
   \label{fig:geo_transfer_target}
\end{figure}

In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
This type of low energy transfer is best suited for autonomous spacecraft missions as the long time of flight is undesirable for manned spacecraft. 
Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
Low energy transfers from the Earth to the Lagrange points are necessary for future missions.

We generate the stable invariant manifold associated with the periodic orbit.
The stable manifold is propagated to the \Poincare section directed along the positive \( \hat{x} \) axis, or \( \alpha =\ang{0} \) denoted by the horizontal black line.
The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section.
The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eqn:constraints} defined on the \Poincare section.
This allows us to approximate the set of states that are achievable in the \( \hat{x} \text{ vs. } \hat{\dot{x}} \) space.
A final state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
With each reachable set we move the final trajectory closer to the stable manifold.
\Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory, with the final trajectory shown in red which ensures the intersection with the stable invariant manifold.
Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_full} 
		\caption{Transfer} \label{fig:geo_transfer_full} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_zoom} 
		\caption{State Trajectory} \label{fig:geo_transfer_zoom} 
	\end{subfigure}

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare} 
		\caption{\Poincare section} \label{fig:geo_transfer_poincare} 
	\end{subfigure} 
	\caption{Geostationary Transfer}
	\label{fig:geo_transfer} 
\end{figure}

>>>>

%---------------------------- EXPLICIT SUMMARY OF CONTRIBUTIONS--------------------------%
In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
Previous results in the design of optimal transfers have relied on suboptimal direct optimization methods and conventional integration techniques.
This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
We derive and solve the Euler-Lagrange equations to generate the necessary conditions for an optimal solution.
This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods.
In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
We avoid the issues inherent in selecting a valid initial condition for optimization.
Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.

@@

% motivation - may need to compute several reachability sets
There are many situations where a more complicated and extensive orbital transfer is desired. 
For example, we present a simulation of transferring from a geostationary orbit to the a periodic orbit about the Moon.
This type of low energy transfer would be most suitable for unmanned spacecraft transitioning from the Earth to the Moon.
The long time of flight would make such a transfer unsuitable for manned missions due to life support constraints.
%This type of low energy transfer is best suited for autonomous spacecraft missions as the long time of flight is undesirable for manned spacecraft. 
Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
Low energy transfers from the Earth to the Lagrange points are necessary for future missions.

The previous example demonstrated the capability of determining an orbital transfer after determining the intersection of the reachability set and a target orbit.
However, it may not be possible to achieve an intersection on the \Poincare section after a single iteration. 
Since the spacecraft has an upper bounded control input and time of flight the reachability set is finite in size. 
As a result, we present a method of performing several iterations of the reachability set computation. 
A straightforward method is presented which allows for a series of reachability sets to be computed which progressively move the trajectory towards the target.
In this manner, it is possible to determine more complicated transfers by a simple selection of states from the reachability set.

% discuss difficulties in applying direct optimization or the invariant manifold method
It is possible to design arbitrary transfers using either a direct optimization or invariant manifold based approach.
The direct optimization method transforms the optimal control problem into a nonlinear programming problem.
Instead of solving the Euler-Lagrange equations the state and control histories are parameterized and solved through any number of mathematical programming methods.
However, due to this parameterization only an approximate solution, which approximates the true optimal solution in the limit, is feasible. 
On the other hand, our method applies an indirect optimization method.
The necessary conditions for optimality are computed and directly solved in generating the reachability set. 
The use of the reachability set also avoids the issues of selecting a valid initial condition.
We select a state on the maximum reachability set which minimizes the distance toward the target. 
This straightforward approach achieves an optimal trajectory and is used to generate general transfers.

The invariant manifold method is difficult to apply to general orbital transfers. 
The manifolds are associated with periodic orbits in the three-body system. 
In this case, an appropriate periodic orbit must first be determined prior to generating the invariant manifold.
Furthermore, there is not guarantee that the invariant manifold will pass through a desired region of space. 
For example, in the Earth-Moon system the unstable manifolds of periodic orbits about \( L_1 \) do not pass close to the Earth, but rather are beyond the geostationary orbit altitude. 
In addition, determining the intersections between various invariant manifolds is not trivial. 
It requires an appropriate \Poincare section and the generation of several invariant manifolds.
There is no clear method of selecting the periodic orbits required based on a the type of transfer desired. 
Determining an intersection between these invariant manifolds generally requires extended flight times, involving several orbits of the primaries, before an appropriate intersection is found.
As a result, it is difficult to generalize this method to arbitrary transfers in the three-body problem.

% describe initial and target states for this example
This numerical simulation demonstrates the ability of using multiple iterations of the reachability set to achieve a more complicated transfer.
In this manner, it is possible to design arbitrary transfers which are not possible using a single reachability computation.
Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional Earth-Moon three-body system. 
It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
The initial and target orbits are illustrated in~\cref{fig:geo_transfer_target}.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
   \caption{Transfer target}
   \label{fig:geo_transfer_target}
\end{figure}
In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
We seek to generate a transfer from the geostationary orbit onto the stable invariant manfiold.
Once on the manifold, the spacecraft will coast in an uncontrolled fashion and asymptotically arrive at the desired periodic orbit.

\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{third_reachability_set} % requires the graphicx package
   \caption{Third reachability set}
   \label{fig:third_reachability_set}
\end{figure}
We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
The stable manifold is propagated to the \Poincare section at \( y = 0 \) denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_max = 0.75 \).
The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eq:constraints} defined on the \Poincare section.
This allows us to approximate the set of states that are achievable in the \( x \text{ vs. } \dot{x} \) space.
After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds.
%A state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
For example,~\cref{fig:third_reachability_set} shows the reachable set in red and the minimum distance to the stable manifold in green.

With each reachable set we move the controlled trajectory closer to the stable manifold.
\Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory, with the final trajectory shown in red which ensures the intersection with the stable invariant manifold.
Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_full} 
		\caption{Transfer} \label{fig:geo_transfer_full} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_zoom} 
		\caption{State Trajectory} \label{fig:geo_transfer_zoom} 
	\end{subfigure}

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare} 
		\caption{\Poincare section} \label{fig:geo_transfer_poincare} 
	\end{subfigure}~%
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{control_input_geo} 
		\caption{Control input} \label{fig:control_input_geo} 
	\end{subfigure} 
	\caption{Geostationary Transfer}
	\label{fig:geo_transfer} 
\end{figure}

This approach demonstrates the ability to link several computations of the reachability set to enable a more general transfer.
We use eight iterations in order to transfer from the geostationary orbit to the stable manifold.
This approach allows for a larger class of potential transfers which leverage the capabilities of low-thrust propulsion systems.
We are able to achieve a transfer that is not possible via the standard invariant manifold approach. 
In addition, this example illustrates a straightforward method to depart from the natural dynamics and transfer to a large region of the phase space that is not accessible via invariant manifolds alone.
